---
title: "CDS HomeWork 2"
author: "Jonathan Rystrøm"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
urlcolor: blue
---

### load packages
```{r}
library(tidyverse)
```


## 1) Use R to figure out how many elements in the vector below are greater than 2.

```{r}
# Define the vector
rooms <- c(1, 2, 1, 3, 1, NA, 3, 1, 3, 2, 1, NA, 1, 8, 3, 1, 4, NA, 1, 3, 1, 2, 1, 7, 1, NA)

# Remove NA's and take the sum of the statemtn
sum(na.omit(rooms) > 2)
```
As seen on the output there are **8** values greater than 2 (excluding the NA's)


## 2) What type of data is in the 'rooms' vector?
This can be answered using the class-function
```{r}
class(rooms)
```
The answer is **numeric**


## 3) What is the result of running median() function on the above 'rooms' vector?
Let's try it out
```{r}
median(rooms)
```
We get an NA. This is because the NA's could conceivably be any value, so R cannot possibly say what the median is.


## 4) Submit an image to Github
Inside your R Project (.Rproj), install the 'tidyverse' package and use the download.file() and read_csv() function to read the SAFI_clean.csv dataset into your R project as 'interviews' digital object (see instructions in https://datacarpentry.org/r-socialsci/setup.html and 'Starting with Data' section). Take a screenshot of your RStudio interface showing a) the script you used to create the object, b) the 'interviews' object in the Environment and the c) structure of your R project in the bottom right Files pane. Save the screenshot as an image and put it in your AUID_lastname_firstname repository inside our Github organisation (github.com/Digital-Methods-HASS) or equivalent. Place here the URL leading to the screenshot in your repository.

The repo can be found [here](https://github.com/Digital-Methods-HASS/au635153_Jonathan_Rystroem/tree/master/HW4)

```{r}
pacman::p_load(tidyverse)
interviews <- read_csv("../data/SAFI_clean.csv")
```


