---
title: "CDS HomeWork 6"
author: "Jonathan Rystrøm"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
urlcolor: blue
---

### load packages
```{r}
pacman::p_load(gapminder, tidyverse)
```

### Loading data
```{r}
dat <- gapminder
```

## Question 1: 
*Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.*

Defensive function means that we should make sure the data input is numeric
```{r}
calculate_gdp <- function(population, gdpPercapita) {
  stopifnot(is.numeric(population) & is.numeric(gdpPercapita))
  return(population * gdpPercapita)
}

dat %>% 
  filter(year %in% c(1967, 1977, 1987, 1997, 2007, 2017), country == "Denmark") %>% 
  mutate(total_gdp = calculate_gdp(pop, gdpPercap)) %>% 
  select(year, total_gdp)
  
```

## Question 2
*Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70. (Hint: use the Control Flow tutorial as a guide here)*
In general, looping in R is slow and generally not recommended as R is a vectorized language. However, here is how to do it. I look at life expectancy in 2017
```{r}
countries <- unique(dat$country)
dat2017 <- filter(dat, year == max(year))

for (land in countries) {
  life_expectancy <- dat2017 %>% 
    filter(country == land) %>% 
    pull(lifeExp)
  
  if (life_expectancy < 50) {
    print(paste0(land, "'s life expectancy is less than 50!"))
  }
  
  else if (between(life_expectancy, 50, 70)) {
    print(paste0(land, "'s life expectancy is between 50 and 70"))
  }
  
  else if (life_expectancy > 70) {
    print(paste0(land, "'s life expectancy is more than 70!"))
  }
}

```


## Question 3
*Challenge/Optional: Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘M’ and graphs life expectancy against time (using plot() function) as a line graph if the mean life expectancy is under 50 years.*
```{r}
for (land in countries) {
  # Find only rows with that country
  country_dat <- dat %>% 
    filter(country == land)
  
  # Get the latest life expectancy
  life_expectancy <- country_dat %>% 
    filter(year == max(year)) %>% 
    pull(lifeExp)
  
  # Checks starts with M and low life expectancy
  if (str_detect(land, "^M") & (life_expectancy < 50)) {
    country_plot <- country_dat %>% 
      ggplot(aes(x=year, y=lifeExp)) + 
      geom_line() + 
      labs(title = paste0("Life Expectancy for ", land)) + 
      theme_minimal()
    plot(country_plot)
  }
}
```

